Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/splitting.tex
Update masking to be consistent with the new 4.0.

@@

To resolve the mismatch between the multi-stage computations
represented by residual tables and partial values and the desired
monostaged outputs of splitting, we choose to define a translation,
called {\em masking}, written $\rtab \xi v \vsplito \mval i q$, which
converts residual tables and partial values into a {\em masked value}.
The first part of the masked value, $i$, is a monostage value holding
the result of all first stage components of $v$.  The second
component, $q$, is a residual representing all second stage
computations in $\xi$ and $v$. (A precise definition of masking is
given in \ref{sec:masking}.)

Masking converts a residual table and associated partial value into a
masked value $\mval i q$.
%that explicitly separates the first stage components of 
%result from the residual.
%where $i$ is a monostage value encoding all
%first stage components of the partial value, and $q$ is a monostage
%term encoding all of the second-stage computations of the partial
%value.
Masking is not a transformation performed as part of the splitting
algorithm.  Instead, masking defines an intuitive notion of the first
stage and second stage ``components'' of residual tables and partial
values, and it is used to define the requirements of what output
splitting of \bbonem\ terms must produce.  A precise definition of the
masking relation is given in \ref{fig:valMask}, but to gain intuition
about the intent of masking operations, consider the following
residual table and partial value, colored as before:
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (next{`2`yhat`1`}, injL (grnd{7}, next{`2`yhat`1`}))`
\end{lstlisting}
To construct a value representing its first-stage components ($i$ in
the resulting masked value), masking first redacts all second stage
(blue) parts, along with the surrounding \texttt{next}
annotations. (This redaction removes the entire residual table since
it only represents stage two computations.) The resulting ``holes'' in
the term are replaced with unit values.
%\begin{lstlisting}
%4`##########`1` (`4`######`1`, injL (grnd{7}, `4`######`1`))`
%\end{lstlisting}
%The holes inside the term needs to be filled with something to convey the lack of information, so we use a unit values:
\begin{lstlisting}
1`((), injL (grnd{7},()))`
\end{lstlisting}
Finally, masking dropping \texttt{grnd} annotations yielding a value.
\begin{lstlisting}
1`((), injL (7,()))`
\end{lstlisting}
To construct a residual corresponding to the stage two computations
represented by the residual table and partial value ($q$ in resulting
the masked value), masking first redacts all \texttt{grnd} blocks
(replaced with unit), \texttt{next} annotations, and injection tags
(essentially, the tuple structure is kept, but all other structure is
discarded):
%\begin{lstlisting}
%2`[yhat |-> 1+2]`1` (`4`####`2`yhat`4`#`1`, `4`####`1` (`4`####`1`,`4`####`2`yhat`4`#`1`))`
%\end{lstlisting}
%Where the \texttt{grnd} blocks left a hole, we put in a unit values:
\begin{lstlisting}
2`[yhat |-> 1+2] (yhat, ((),yhat))`
\end{lstlisting}
Then masking reifies the residual table into \texttt{let} bindings,
yielding a residual:
\begin{lstlisting}
2`let yhat = 1+2 in (yhat, ((),yhat))`
\end{lstlisting}

Note in \ref{fig:valMask} that since lambdas may represent multistage
computations, masking splits the body of lambdas as general world
\bbonem\ terms (as described in \ref{sec:split-one}), and packages the
resulting terms as functions in both the first and second stage parts
of the masked value.
% Masking operates by first inducting on the entries of the residual table.  
% Being purely second-stage content, these are reified into let statements at the top of the resumer.
% Once the table is empty, masking inducts on value itself.

>>>>

\newpage

@@

%Masking separates residual tables and partial values into first stage
%results and second stage residuals.

Masking (given by the relation $\vsplito$) converts a residual table $\xi$
and associated partial value $v$ into a masked value $\mval i q$.
%that explicitly separates the first stage components of 
%result from the residual.
%where $i$ is a monostage value encoding all
%first stage components of the partial value, and $q$ is a monostage
%term encoding all of the second-stage computations of the partial
%value.
The first part of the masked value, $i$, is a monostage value holding
the result of all first stage components of $v$.  The second
component, $q$, is a residual representing all second stage
computations in $\xi$ and $v$.

%(A precise definition of masking is
%given in \ref{sec:masking}.)

%Masking is not a transformation performed as part of the splitting
%algorithm.  Instead, masking defines an intuitive notion of the first
%stage and second stage ``components'' of residual tables and partial
%values, and it is used to define the requirements of what output
%splitting of \bbonem\ terms must produce.
To provide intuition about the intent of masking, consider the
following residual table and partial value:
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (next{`2`yhat`1`}, injL (grnd{7}, next{`2`yhat`1`}))`
\end{lstlisting}
To construct the value $i$ representing its first-stage components,
masking first redacts all second stage (blue) parts, along with the
surrounding \texttt{next} annotations. (This redaction removes the
entire residual table since it only represents second stage
computations.) The resulting ``holes'' in the term are replaced with
unit values.
%\begin{lstlisting}
%4`##########`1` (`4`######`1`, injL (grnd{7}, `4`######`1`))`
%\end{lstlisting}
%The holes inside the term needs to be filled with something to convey the lack of information, so we use a unit values:
\begin{lstlisting}
1`((), injL (grnd{7},()))`
\end{lstlisting}
Finally, masking drops \texttt{grnd} annotations yielding:
\begin{lstlisting}
1`((), injL (7,()))`
\end{lstlisting}
To construct the residual $q$ (corresponding to second stage computations) masking first redacts all \texttt{grnd} blocks
(replacing them with unit), \texttt{next} annotations, and injection tags:
%\begin{lstlisting}
%2`[yhat |-> 1+2]`1` (`4`####`2`yhat`4`#`1`, `4`####`1` (`4`####`1`,`4`####`2`yhat`4`#`1`))`
%\end{lstlisting}
%Where the \texttt{grnd} blocks left a hole, we put in a unit values:
\begin{lstlisting}
2`[yhat |-> 1+2] (yhat, ((),yhat))`
\end{lstlisting}
Then the residual table is reified into \texttt{let} bindings,
yielding:
\begin{lstlisting}
2`let yhat = 1+2 in (yhat, ((),yhat))`
\end{lstlisting}
A precise definition of the masking relation is given in
\ref{fig:valMask}.

%Since lambdas may represent multi-stage computations, masking splits
%the body of lambdas as general world \bbonem\ terms (as described in
%\ref{sec:split-one}), and packages the resulting terms as functions in
%both the first and second stage parts of the masked value.

% Masking operates by first inducting on the entries of the residual table.  
% Being purely second-stage content, these are reified into let statements at the top of the resumer.
% Once the table is empty, masking inducts on value itself.
